---
editor: source
toc: true
title-block-banner: "#0057b8"
subtitle: Cohort 3.0
echo: false
warning: false
---

# Module 2: How do we build data solutions?

You are building a donor churn prediction model. You are reviewing the datasets utilized to build the donor churn model in order to develop an ETL (Extract-Transform-Load) pipeline.

Review the raw datasets that are utilized to generate each element of the donor churn model. Consider the questions about data transformation. These include, but are not limited to:

* Are there any issues with the data (outliers, missing, etc.)?
* Is the data validated?
* Does the data need to be aggregated? How?
* Do we need to augment the data?
* Does it need to be joined or merged? How?


What would we need to do to transform the raw data into the data utilized for the analysis?

## Dataset After Transformation
After all datasets are transformed, there will be one record per person per year. It will look something like this:

```{r}
library(tidyverse)
library(DT)

tribble(
  ~Person,       ~`Donation $ Last Year`, ~`Event Participation Last Year`, ~`Email Clickthrough Rate Last Year`, ~`Income Last Year`, ~`Donated This Year`,
  "Mickey Mouse",                    200,                                0,                                 0.05,              200000,                 TRUE,
  "Bugs Bunny",                       50,                                3,                                 0.07,               40000,                FALSE,
  "Patrick Star",                     25,                                1,                                 0.03,               32000,                 TRUE,
  "Bruce Wayne",                    1000,                                0,                                 0.01,           100000000,                FALSE,
  "Peter Parker",                     30,                                5,                                 0.21,               20000,                 TRUE,
  "Ash Ketchum",                      75,                                2,                                 0.06,              120000,                   NA,
  "Velma Dinkley",                   100,                                5,                                 0.12,               67000,                   NA
) %>% 
  datatable(
    rownames = FALSE,
    extensions = 'Buttons',           # load Buttons
    options = list(
      pageLength = 10,
      autoWidth = TRUE,
      searchHighlight = TRUE,
      dom       = 'Bfrtip',           # B = buttons, f = filter, r = processing, t = table, i = info, p = pagination
      buttons   = list(
        list(extend   = 'csv',
             filename = 'module_2_activity-data_after_transformations'),
        list(extend   = 'excel',
             filename = 'module_2_activity-data_after_transformations')
      )
    )
  ) %>%
  formatCurrency("Donation $ Last Year", currency = "$", digits = 0) %>%
  formatCurrency("Income Last Year", currency = "$", digits = 0) %>%
  formatPercentage("Email Clickthrough Rate Last Year", digits = 1)
```

## Raw Datasets
Each of the following four datasets represents the raw data utilized to create the table above.

::: {.panel-tabset}

### Past Donation History
This dataset originates from **Salesforce**. It is updated each time a donation is made.

```{r}
read_csv("donations_synthetic_salesforce_v3.csv") %>% 
  mutate(`Date of Donation` = as.Date(`Date of Donation`, format = "%m/%d/%y")) %>% 
  datatable(
    rownames = FALSE,
    extensions = 'Buttons',           # load Buttons
    options = list(
      pageLength = 10,
      autoWidth = TRUE,
      searchHighlight = TRUE,
      dom       = 'Bfrtip',           # B = buttons, f = filter, r = processing, t = table, i = info, p = pagination
      buttons   = list(
        list(extend   = 'csv',
             filename = 'module_2_activity-past_donation_history'),
        list(extend   = 'excel',
             filename = 'module_2_activity-past_donation_history')
      )
    )
  ) %>%
  formatCurrency("Dollar Amount", currency = "$", digits = 2)
```

### Event Attendance
This dataset originates from a **Google Form**. It is updated during each event hosted by the organization.

```{r}
read_csv("event_tracker_attendance_sheet.csv") %>% 
  datatable(
    rownames = FALSE,
    extensions = 'Buttons',           # load Buttons
    options = list(
      pageLength = 10,
      autoWidth = TRUE,
      searchHighlight = TRUE,
      dom       = 'Bfrtip',           # B = buttons, f = filter, r = processing, t = table, i = info, p = pagination
      buttons   = list(
        list(extend   = 'csv',
             filename = 'module_2_activity-event_attendance'),
        list(extend   = 'excel',
             filename = 'module_2_activity-event_attendance')
      )
    )
  )
```

### Engagement with Emails
This dataset originates from **Hubspot**. It is automatically updated as individuals on the email list interact with emails sent by the organization.

```{r}
read_csv("hubspot_email_interactions.csv") %>% 
  datatable(
    rownames = FALSE,
    extensions = 'Buttons',           # load Buttons
    options = list(
      pageLength = 10,
      autoWidth = TRUE,
      searchHighlight = TRUE,
      dom       = 'Bfrtip',           # B = buttons, f = filter, r = processing, t = table, i = info, p = pagination
      buttons   = list(
        list(extend   = 'csv',
             filename = 'module_2_engagement_with_emails'),
        list(extend   = 'excel',
             filename = 'module_2_engagement_with_emails')
      )
    )
  )
```

### Income Data
This dataset lives in **Salesforce**. A development manager updates the data about once per year.

```{r}
read_csv("donor_volunteer_master_v2.csv") %>% 
  datatable(
    rownames = FALSE,
    extensions = 'Buttons',           # load Buttons
    options = list(
      pageLength = 10,
      autoWidth = TRUE,
      searchHighlight = TRUE,
      dom       = 'Bfrtip',           # B = buttons, f = filter, r = processing, t = table, i = info, p = pagination
      buttons   = list(
        list(extend   = 'csv',
             filename = 'module_2_activity-income_data'),
        list(extend   = 'excel',
             filename = 'module_2_activity-income_data')
      )
    )
  ) %>%
  formatCurrency("Net Worth", currency = "$", digits = 2)
```