---
editor: source
toc: true
title-block-banner: "#0057b8"
subtitle: Cohort 3.0
echo: false
warning: false
---

# Module 3: How do we tell data-driven stories?

You are developing a presentation for members of your state legislature outlining childhood obesity rates across ethnicities in your state. You are considering data for the last 10 years in your presentation.

Consider how you would tell the story with data to this group of policy makers.

## Activity 1

### Activity Objective
Consider the components of a data story. How does this scenario map to those components?
What narrative structures might you consider utilizing? Why?

### Elements of data stories

Remember the elements of data stories that you might want to utilize. Begin with the end in mind.

#### Audience
Who are you speaking to?

* Are they internal? External?
* Are they an expert? A novice?
* What do they care about?
* Why are they there?

#### Give Context to Issues
Consider utilizing these types of narrative structures:

* Narrate change over time
* Start big and drill down
* Highlight contrasts
* Profile the outliers
* Start small and zoom out
* Dissect the factors
* Explore the intersection

#### Solution
Begin with the end in mind. Consider starting by understanding your solution:

* What should the audience takeaway?
* Is there a specific call-to-action?
* What behaviors should change?
* What is the expected result from the change?


## Activity 2

```{r}
#| echo: false
#| warning: false

if(! require(pacman)) install.packages("pacman")

pacman::p_load(
  tidyverse,
  DT
)
```

### Sample data
The data included below represents a potential sample of data for you to consider. Note that this data has been generated for the purposes of this exercise and does not represent real world data.

```{r}
#| echo: false

datatable(
  tidyr::tribble(
    ~Year, ~`White (%)`, ~`Black or African (%)`, ~`Asian (%)`, ~`Hispanic (%)`,
    2015,  20.0,         26.0,                    24.0,         21.0,
    2016,  16.0,         29.0,                    23.0,         22.0,
    2017,  18.0,         27.0,                    22.0,         23.0,
    2018,  17.0,         29.0,                    21.0,         24.0,
    2019,  18.0,         26.0,                    20.0,         25.0,
    2020,  24.0,         21.0,                    20.0,         25.0,
    2021,  17.0,         22.0,                    22.5,         22.5,
    2022,  22.0,         25.0,                    25.0,         20.0,
    2023,  24.0,         20.0,                    27.5,         17.5,
    2024,  17.0,         27.0,                    30.0,         15.0
  )
)
```


### Activity Objective
Sketch three potential visualizations you might utilize to tell this story.
Consider which types of visualizations might be useful for you.
Consider how the visualizations fit into the broader data story you are trying to tell.

Make your sketch wherever you're comfortable. Be prepared to share your sketch with group mates. Some recommendations:

* Pen and paper!
* Presentation tools like PowerPoint or Google Presentations
* Online whiteboarding tools like [Miro](https://miro.com/) or [Vibe](https://app.vibe.us)
* [Virtual Graph Paper](https://virtual-graph-paper.com/)


## Activity 3

```{r}
#| echo: false
#| warning: false

if(! require(pacman)) install.packages("pacman")

pacman::p_load(
  tidyverse,
  here,
  DT
)
```

Consider Grade 3 reading proficiency by race in Delaware’s school districts from 2015 to 2021.

Compose a data visualization for a policy maker, school administrator, or member of the public that encapsulates the best practices discussed throughout today’s presentation.

Data is sourced from [Delaware’s Open Data Portal](https://data.delaware.gov/Education/Student-Assessment-Performance/ms6b-mt82/about_data).

### Data

#### Data Dictionary
The following fields are included in the dataset:

| Field           | Description                                                                                                                                                                                                                                                                                                                      |
|-----------------|----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| `School Year`     | School year for which record is applicable. For example, 2016 = school year which ended in June 2016.                                                                                                                                                                                                                            |
| `District Code`   | Number representing each School District. Statewide data rows will have “0” in this column.                                                                                                                                                                                                                                      |
| `District`        | Full name of the School District. Statewide data rows have "State of Delaware" in this column.                                                                                                                                                                                                                                   |
| `School Code`     | Number representing each School within the school district. Statewide or Districtwide data rows will have “0” in this column.                                                                                                                                                                                                    |
| `Organization`    | Full name of the Organization which is the School if School Code is given in the row. Districtwide data rows give full name of School District. Statewide rows give "State of Delaware" in this column.                                                                                                                          |
| `Assessment Name` | Statewide summative assessment name. Valid values include: ACCESS – EL Annual ACCESS assessment; DCAS – Delaware Comprehensive Assessment System; DCAS-Alt1 – Delaware Comprehensive Assessment System – Alternate Assessment; DeSSAALT – Delaware System; SAT11 – SAT assessment; SBAC – Smarter Balanced Assessment Consortium |
| `Content Area`    | Statewide summative assessment subject (or content area). Valid values include: CMP – EL Annual ACCESS Composite Score; ELA – English Language Arts/Reading; ESSAY – Essay Component of SAT assessment; MATH – Mathematics; SCI – Science; SOC – Social Studies;                                                                 |
| `Race`            | Represents the race/ethnicity of the unique group of students within a school/district.                                                                                                                                                                                                                                          |
| `Gender`          | Represents the grade level of the unique group of students within a school/district.                                                                                                                                                                                                                                             |
| `Grade`           | Represents the grade level of the unique group of students within a school/district.                                                                                                                                                                                                                                             |
| `SpecialDemo`     | Represents the special population status of the unique group of students within a school/district.                                                                                                                                                                                                                               |
| `Geography`       | Represents the geography of the unique group of students within a school/district. Geography can have the following values: ALL – all geographies combined                                                                                                                                                                       |
| `SubGroup`        | Names the unique group of students within a school/district described by the combination of Race, Gender, Grade, SpecialDemo, and Geography.                                                                                                                                                                                     |
| `RowStatus`       | Indicates whether the aggregate data in this row has either been REDACTED or REPORTED. If redacted, certain data has been hidden to comply with state and federal privacy laws. For more information, contact the Delaware Department of Education.                                                                              |
| `Tested`          | The number of students who completed the statewide summative assessment listed with a valid score                                                                                                                                                                                                                                |
| `Proficient`      | The number of students who earned a proficient achievement level. This is typically considered as 3 or above.                                                                                                                                                                                                                    |
| `PctProficient`   | The percent of students who earned an achievement level considered proficient. This is the number of students who earned a proficient achievement level divided by the number of students who tested.                                                                                                                            |
| `ScaleScoreAvg`   | The average scale score for the specified group of students.                                                                                                                                                                                                                                                                     |

#### Sample

You can view a selection of the data below. Sample data includes the following filters:

* `Gender` is filtered to show `All Students`
* `School Code` is filtered to `0`, which forces the `Organization` to be at the district or charter school level
* `SpecialDemo` is filtered to show `All Students`

```{r}
#| echo: false
#| warning: false

reading_level <- read_csv(here("Student_Assessment_Performance_20241020.csv"))

```

```{r}
#| echo: false
reading_level %>% 
  filter(Gender == "All Students",
         `School Code` == 0,
         SpecialDemo == "All Students") %>% 
  datatable(
    rownames = FALSE,
    extensions = 'Buttons',           # load Buttons
    options = list(
      pageLength = 10,
      autoWidth = TRUE,
      searchHighlight = TRUE,
      dom       = 'Bfrtip',           # B = buttons, f = filter, r = processing, t = table, i = info, p = pagination
      buttons   = list(
        list(extend   = 'csv',
             filename = 'activity_3-grade3_reading'),
        list(extend   = 'excel',
             filename = 'activity_3-grade3_reading')
      )
    )
  )
```


### Activity Objective
Develop a data narrative that you want to tell based upon Grade 3 reading proficiency data in Delaware.
Consider all of the elements of a data narrative.
Consider which types of graphs would be most effective to tell your story.
Consider how you would incorporate best practices into those graphs.

Sketch your narrative including at least three graphs.

